---
title: "Calculate DBH from iPhone LiDAR"
author: "Hauke Gronenberg"
date: "4/12/2021"
output: html_document
---

```{r setup, echo=F, results='hide', warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
library(lidR)
library(tidyverse)
library(data.table)
```

In the following, I will present a first draft on what can be done with the iPhone's 
LiDAR sensor in the field of forestry. My goal is to calculate the 
[diameter at breast height (DBH)](https://en.wikipedia.org/wiki/Diameter_at_breast_height).

The R code and the data of this analysis are available as 
[GitHub repository](https://github.com/haukegron/iLiDAR).

## Data Collection

For the scan I used the iOS App 
[3d Scanner App](https://apps.apple.com/us/app/3d-scanner-app/id1419913995), 
since it allows to export the data as [LAS file](https://en.wikipedia.org/wiki/LAS_file_format).

## Data Analysis

Most of the analysis was done with the great R package [lidR](https://github.com/Jean-Romain/lidR).

### Raw Point Cloud

This is a plot of the unprocessed point cloud. Thanks to the combination of the
iPhone's LiDAR sensor and the camera we have the color information for each point.

```{r warning=FALSE, message=FALSE, echo = F}
las <- lidR::readLAS(system.file("extdata","trees_on_hill.laz",package = "iLiDAR"))

# fix wrong orientation
z <- las@data$Z

las@data$Z <- las@data$Y
las@data$Y <- z
```

```{r, webgl=TRUE, echo = F}
lidR::plot(las, color = "RGB", axis=T)
```

### Ground Classification and Height Normalization

For the ground classification of the point cloud lidR's build-in 
*Cloth Simulation Function* (based on 
[Zhang et al. 2016](https://www.mdpi.com/2072-4292/8/6/501/htm)) was used. 
For further information on ground classification have look at section 
[3 Ground classification](https://jean-romain.github.io/lidRbook/gnd.html) 
of the lidR-book .
In the following, the `lidR::tin()`-function ([R help](https://rdrr.io/cran/lidR/man/tin.html)) 
was used to normalize the point cloud.

```{r ground_classification}
las <- classify_ground(las, algorithm = csf(sloop_smooth = T))

las <- lidR::normalize_height(las       = las,
                              algorithm = lidR::tin(),
                              use_class = c(2))
```

### Tree Segmentation

The next step of the analysis is to identify clusters in the point cloud.
Therefore, the function `lidR::watershed()` 
([R help](https://rdrr.io/cran/lidR/man/watershed.html)) was used.
Due to the iPhone's LiDAR-sensor limitations and the chosen range setting within
the [3d Scanner App](https://apps.apple.com/us/app/3d-scanner-app/id1419913995) 
only a few points do occur above 3 meters. All of them will be excluded from 
further analysis.
Additionally, I am not interested in points below 1.2 meters.

```{r}
las <- lidR::filter_poi(las, Classification != 2)

las <- filter_poi(las, Z < 3,
                       Z > 1.2)

chm <- grid_canopy(las, 0.1, p2r())

algo <- watershed(chm, th_tree = 1.2, tol = 1, ext = 1)
las <- segment_trees(las, algo)

```

As we can see, the segmentation was applied successfully.
However, I have to get rid of the bushes since I am only interested on trees.

```{r, webgl=TRUE, echo = F}
plot(las, color = "treeID")
```

### Tree Hulls and Selection of *Trees* by Tree Tops

For each segmented tree (including the remaining bushes), the convex hull was calculated with the function `lidR::delineate_crowns` ([R help](https://rdrr.io/cran/lidR/man/delineate_crowns.html)).
To identify trees (and subsequently remove bushes) the build-in local maximum filter `lmf()` ([R help](https://rdrr.io/cran/lidR/man/lmf.html)) was applied.


```{r}
hulls <- lidR::delineate_crowns(las,
                       type = "convex",
                       attribute = "treeID")
# find trunk location
ttops <- find_trees(las, lmf(ws = 1))
```

```{r, echo=F}
plot(hulls)
plot(ttops, add = T)
```

```{r}
sel <- sp::over(x = ttops, y = hulls)

las <- lidR::filter_poi(las, treeID %in% sel$treeID)
```

### Calculation of the DBH

The final step of this analysis is to calculate the [diameter at breast height](https://en.wikipedia.org/wiki/Diameter_at_breast_height) for all identified trees.
Therefore, I made use of code developed by [Leon](https://github.com/Lenostatos).

```{r , warning=F}
las_dbh <- filter_poi(las, Z < 1.31,
                           Z > 1.29)

ellipsoids <- iLiDAR::create_crown_ellipsoids(las_dbh)
```

```{r, echo = F, warning=F}

# source: https://stackoverflow.com/questions/6528180/ggplot2-plot-without-axes-legends-etc
t <- theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())

  # source: https://yutani.rbind.io/post/geom-sf-text-and-geom-sf-label-are-coming/
  # label
  pos <- sf::st_point_on_surface(ellipsoids)

  pos <- as.data.frame(sf::st_coordinates(pos))
  pos$Y <- pos$Y - 0.5
  pos$dbh <- round(ellipsoids$diameter_ellipsoid_mean,2)
  
  p <- ggplot(ellipsoids) +
       geom_sf()
  
  p + geom_text(data = pos, aes(X, Y, label = dbh), colour = "black") +t
  
```
